---
title: "Tidy Tuesday 2021 Week 13"
author: "Sean Jackson"
date: "3/26/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
## initial settings
knitr::opts_chunk$set(
  comment = NA,
  echo = TRUE,
  error = TRUE,
  cache = FALSE,
  message = FALSE,
  dpi = 96,
  warning = FALSE
)

## width to use when printing tables etc.
options(
  width = 250,
  scipen = 100,
  max.print = 5000,
  stringsAsFactors = FALSE
)
```

<style>
.table {
  width: auto;
}
ul, ol {
  padding-left: 18px;
}
pre, code, pre code {
  overflow: auto;
  white-space: pre;
  word-wrap: normal;
  background-color: #ffffff;
}
</style>

United Nations Roll Call votes is a wealth of interesting data. This week I will take the short names of each roll call vote and turn them into yearly wordclouds with color and size representing word frequency (blue low and red high). First I load the libraries, source data, and an image.

```{r libraries and data}
library(tidyverse)       # Data management
library(ggplot2)     # Graph framework
library(wordcloud)
library(ggwordcloud) # Wordcloud using ggplot
library(gganimate)   # Animation
library(tm)
library(RColorBrewer) # color palettes
library(tidytext)
library(lubridate)
library(png)
library(grid)

roll_calls <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-23/roll_calls.csv')
img <- readPNG("unv3.png")
```

Next I perform a few alterations to the data to create a year column, break up the strings in short into individual words, remove common words, count the distinct words each year, finally and filter out the lower word counts.

```{r clean data}
rc_words <- roll_calls %>%
  mutate(yr = as.integer(year(date))) %>%
  filter(!is.na(short)) %>%
  unnest_tokens(word, short) %>%
  anti_join(stop_words, by = "word") %>%
  distinct(yr, word) %>%
  add_count(word, name = "word_count") %>%
  filter(word_count >= 25)
```

Now I use ggplot and geom_text_wordcloud to build the base frames out of the data and image then animate the gif. 

```{r format and animate}
gg <- rc_words %>%
  ggplot(.,aes(label = word
             , size=word_count
             , color = word_count)) +  
  annotation_custom(rasterGrob(img, 
                               width = unit(1,"npc"), 
                               height = unit(1,"npc")), 
                    -Inf, Inf, -Inf, Inf) +
  geom_text_wordcloud() +
  scale_size_area(max_size = 7) +
  scale_colour_gradient(low =  "blue", high = "red") +
  labs(title = "Annual Wordcloud of UN Roll Call Votes",
       caption = "Data from United Nations\nSeanPJ.com") +
  theme_minimal() +  
  theme(plot.title = element_text(hjust = 0.5, face="bold", size=15),
        plot.subtitle = element_text(hjust = 0.5)) +
  labs(subtitle = '{frame_time}') +
  transition_time(yr)

animate(gg, nframes = 120, fps = .5,  width = 600, height = 400,end_pause=30)
```

